---
title: "Indica offset"
output: html_notebook
---

```{r}
library(tmap)
library(tidyverse)
library(LEA)
library(vegan)
library(sf)
```

```{r}
indica_env_variables = readRDS("indica_env_variables.rds")

ggplot(indica_env_variables,aes(y = value, 
                     x = ssp,
                     color = ssp)) +
  geom_point() +
  facet_grid(~name) +
  xlab(NULL) +
  theme(axis.text.x = element_blank())
```
Looking at geographical distribution of geotagged indica accessions

```{r}

inds_geotagged = readRDS("indica_geotagged.rds")

data(World)
asia = World %>% dplyr::filter(continent == "Asia")

ggplot(asia) +
  geom_sf() +
  geom_point(inds_geotagged,mapping = aes(x = LON,y = LAT))

```
I am working with the dataset from the Gutaker et al. 2020 study.

The dataset has been filtered to retain biallelic SNPs with genotyping rate > 0.8, at HWE for F = 0.96, and then imputed at missing genotypes.

For population structure analysis, used a linkage-pruned set.

Plotting eigenvalues of prcomp run on genotype matrix (don't run)


```{r}
#ind_pca = readRDS("ind_prcomp.rds")

plot(ind_pca, col = "blue")
```
This was used to run admixture analysis for indica genotypes. Don't run.

```{r}
ind_project = NULL
ind_project = snmf("INDICA/indica_10kb_50_filtered_maf0.05.geno",
K = 4:11,
entropy = TRUE,
repetitions = 10,
project = "new")
```

Cross-entropy across different values of K

```{r}
ind_project = readRDS("indica_structure.rds")

plot(ind_project, col = "blue", pch = 19, cex = 1.2)
```

It looks a little messier than japonica, but going with K = 7 based partly on the Gutaker paper

```{r}

K = 7

ind_best = which.min(cross.entropy(ind_project, K = K))

ind_admix = as.data.frame(Q(ind_project , K = K, run = ind_best))
ind_admix$ID = inds_geotagged$ID

ind_admix = left_join(ind_admix,inds_geotagged) 

ind_admix = ind_admix %>% dplyr::arrange(INDK7)

ind_admix_long = tidyr::pivot_longer(ind_admix,cols=1:7)
ind_admix_long = ind_admix_long %>% rowwise() %>%
  dplyr::mutate(ord_ID = paste0(ID,"_",name))


ggplot(ind_admix_long,aes(x=factor(ID),y = value,color = factor(name))) +
  geom_col() +
  facet_grid(~INDK7,space = "free",scales = "free") +
  theme(axis.text.x = element_blank())

```

Before we run the actual GEA, looking at the correlatedness between the three environmental variables:


```{r}

indica_env_list = readRDS("indica_env_list.rds")

cor(indica_env_list[["present"]][,4:6])

```

Performing PCA

```{r}
indica_pred.pca <- vegan::rda(indica_env_list[["present"]][,4:6], scale=T)

summary(indica_pred.pca)$cont
```
```{r}
screeplot(indica_pred.pca, main = "Screeplot: Eigenvalues of Predictor Variables")

```

PC1 seems largely explanatory, using that as the environmental variable for GEA

```{r}
indica_env.pc1 = vegan::scores(indica_pred.pca, choices=1, display="sites", scaling=0)
#saveRDS(env.pc1,"env.PC1.rds")
```

Looking at correlation between PC1 and the environmental variables

```{r}
cor(data.frame(indica_env_list[["present"]][,4:6],PC1 = indica_env.pc1))
```
```{r}
#ind_lf_matrix = data.table::fread("indica_filtered_maf0.05.lfmm")

#ind_lf_PC_K7 = lfmm2(input = ind_lf_matrix, env = ind_env.pc1,
 #     K = 7)
#ind_p_PC_K7 = lfmm2.test(ind_lf_PC_K7,ind_lf_matrix,ind_env.pc1)

```

Looking at distribution of p-values

```{r}

ind_p_PC_K7 = readRDS("ind_p_PC1_K7.rds")

hist(ind_p_PC_K7$pvalues)

```
qqplot is a bit inflated, but does imply presence of outliers:

```{r}

qqplot(rexp(length(ind_p_PC_K7$pvalues), rate = log(10)),
       -log10(ind_p_PC_K7$pvalues), xlab = "Expected quantile",
       pch = 19, cex = .4)
abline(0,1)
```



Importing SNP positions:

```{r}
indica_snp_table = read.table("indica_filtered_maf0.05.map")

indica_snp_table$p = ind_p_PC_K7$pvalues

indica_snp_table$index = 1:length(indica_snp_table$p)
```

Correcting for false-positives and making a Manhattan plot

```{r}

ind_fdr_K7 = p.adjust(ind_p_PC_K7$pvalues,"BY",length(ind_p_PC_K7$pvalues))

sum(ind_fdr_K7 < 0.1)

indica_snp_table$fp = ind_fdr_K7 

indica_snp_table = indica_snp_table %>% dplyr::mutate(sig = case_when(fp < 0.1 ~ "sig.",
                                            fp >= 0.1 ~ "non-sig."
))


ggplot(indica_snp_table,aes(x = index, y = -log10(p),color=sig)) +
  geom_point()


```


A lot of the SNPs are clearly linked. Picking the most significant SNPs from every contiguous series of SNPs (within 100kb) and making another Manhattan plot:


```{r}

ind_snps_K7 = indica_snp_table %>% dplyr::filter(fp < 0.1)

ind_chain = rep(1,length(ind_snps_K7$V4))

for (i in 2:length(ind_snps_K7$V4)){
  if (ind_snps_K7$V1[i] != ind_snps_K7$V1[i-1] | 
      ind_snps_K7$V4[i] - ind_snps_K7$V4[i-1] > 100000){
    ind_chain[i:length(ind_chain)] = ind_chain[i] + 1
  }
}

ind_snps_K7$chain = ind_chain

ind_snps_K7_loci = ind_snps_K7 %>% dplyr::group_by(chain) %>%
  dplyr::filter(p == min(p)) %>%
  dplyr::filter(V4 == min(V4))



#Show these loci on Manhattan plot:

ggplot(indica_snp_table,aes(x = index, y = -log10(p))) +
geom_point(color = "grey") +
geom_point(data = ind_snps_K7_loci, mapping = 
             aes(x = index, y = -log10(p)),color = "blue")


```
Genetic offset was calculated using the following code (don't run)

```{r}

ind_offset_2.6 = genetic.gap(input = ind_lf_matrix,
            env = indica_env_list[["present"]][,4:6],
            pred.env = indica_env_list[["SSP1-2.6"]][,4:6],
            candidate.loci = ind_snps_K7_loci$index,
            K = 7
            )

ind_offset_7.0 = genetic.gap(input = ind_lf_matrix,
            env = indica_env_list[["present"]][,4:6],
            pred.env = indica_env_list[["SSP3-7.0"]][,4:6],
            candidate.loci = ind_snps_K7_loci$index,
            K = 7
            )

ind_offset_8.5 = genetic.gap(input = ind_lf_matrix,
            env = indica_env_list[["present"]][,4:6],
            pred.env = indica_env_list[["SSP5-8.5"]][,4:6],
            candidate.loci = ind_snps_K7_loci$index,
            K = 7
            )


ind_offset_past = genetic.gap(input = ind_lf_matrix,
            env = indica_env_list[["present"]][,4:6],
            pred.env = indica_env_list[["past"]][,4:6],
            candidate.loci = ind_snps_K7_loci$index,
            K = 7
            )
```

Importing these objects:

```{r}
ind_offset_2.6 = readRDS("ind_offset_2.6.rds")
ind_offset_7.0 = readRDS("ind_offset_7.0.rds")
ind_offset_8.5 = readRDS("ind_offset_8.5.rds")
ind_offset_past = readRDS("ind_offset_past.rds")

```

```{r}

inds = read.table("inds_order.txt")

inds_offset = data.frame(inds$V1,
                  ind_offset_2.6$offset,ind_offset_2.6$distance,
                  ind_offset_7.0$offset,ind_offset_7.0$distance,
                  ind_offset_8.5$offset,ind_offset_8.5$distance,
                  ind_offset_past$offset,ind_offset_past$distance)


colnames(inds_offset) = c("ID",
                   "offset_2.6","rona_2.6",
                   "offset_7.0","rona_7.0",
                   "offset_8.5","rona_8.5",
                   "offset_past","rona_past"
                   )



inds_offset = dplyr::left_join(inds_offset,inds_geotagged)
```
Looking at the maximum offset value for plotting range

```{r}
max(inds_offset$offset_2.6,
    inds_offset$offset_7.0,
    inds_offset$offset_8.5,
    inds_offset$offset_past)
```

```{r}
ind_plot_off_2.6 = ggplot(asia) +
  geom_sf() +
  geom_point(data = inds_offset, 
             mapping = aes(x = LON,y = LAT, color = offset_2.6)) +
            scale_colour_gradientn( colors = c("lightblue1","darkblue"),limits=c(0,0.1)) +
   theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_blank(),
        legend.position = "none",
        axis.ticks = element_blank()) +
  ggtitle("Model 2.6")
  
```

```{r}
ind_plot_off_7.0 = ggplot(asia) +
  geom_sf() +
  geom_point(data = inds_offset, 
             mapping = aes(x = LON,y = LAT, color = offset_7.0)) +
            scale_colour_gradientn( colors = c("lightblue1","darkblue"),limits=c(0,0.1)) +
   theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_blank(),
        legend.position = "none",
        axis.ticks = element_blank()) +
  ggtitle("Model 7.0")
  
```


```{r}

ind_plot_off_8.5 = ggplot(asia) +
  geom_sf() +
  geom_point(data = inds_offset, 
             mapping = aes(x = LON,y = LAT, color = offset_8.5)) +
   scale_colour_gradientn( colors = c("lightblue1","darkblue"),limits=c(0,0.1)) +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_blank(),
        legend.position = "none",
        axis.ticks = element_blank()) +
  ggtitle("Model 8.5")
```


```{r}
ind_plot_off_past = ggplot(asia) +
  geom_sf() +
  geom_point(data = inds_offset, 
             mapping = aes(x = LON,y = LAT, color = offset_past)) +
            scale_colour_gradientn( colors = c("lightblue1","darkblue"),limits=c(0,0.1)) +
   theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_blank(),
        legend.position = "none",
        axis.ticks = element_blank()) +
  ggtitle("Past")
  
```


```{r}

ind_plot_with_legend = ggplot(asia) +
  geom_sf() +
  geom_point(data = inds_offset, 
             mapping = aes(x = LON,y = LAT, color = offset_8.5)) +
   scale_colour_gradientn( "Offset",colors = c("lightblue1","darkblue"),limits=c(0,0.1)) +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_blank(),
        axis.ticks = element_blank()) +
  ggtitle("Model 8.5")

ind_legend <- cowplot::get_legend(
  ind_plot_with_legend +  
      theme(legend.position = "bottom") 
)
```


```{r}
ind_prow = cowplot::plot_grid(ind_plot_off_2.6,ind_plot_off_7.0,
                          ind_plot_off_8.5,ind_plot_off_past)

cowplot::plot_grid(ind_prow,ind_legend,ncol = 1,rel_heights = c(7,1))
```


