---
title: "Japonica offset"
output: html_notebook
---

```{r}
library(tmap)
library(tidyverse)
library(LEA)
library(vegan)
library(sf)
```


Plotting environmental variables

```{r}

env_variables = readRDS("env_variables.rds")

ggplot(env_variables,aes(y = value, 
                     x = ssp,
                     color = ssp)) +
  geom_point() +
  facet_grid(~name) +
  xlab(NULL) +
  theme(axis.text.x = element_blank())

```

Looking at geographical distribution of geotagged japonica accessions

```{r}


japs_geotagged = readRDS("japonica_geotagged.rds")

data(World)
asia = World %>% dplyr::filter(continent == "Asia")

ggplot(asia) +
  geom_sf() +
  geom_point(japs_geotagged,mapping = aes(x = LON,y = LAT))

```

I am working with the dataset from the Taiwan study (Alam et al. 2021).

The dataset has been filtered to retain biallelic SNPs with genotyping rate > 0.8, at HWE for F = 0.96, and then imputed at missing genotypes.

For population structure analysis, used a linkage-pruned set.

Plotting eigenvalues of prcomp run on genotype matrix (don't run)

```{r}

#jap_pca = readRDS("jap_prcomp.rds")

plot(jap_pca, col = "blue")
```
Ran population structure analysis using the following code:

```{r}
project = NULL
project = snmf("japonica_10kb_50_filtered_maf0.05.geno",
K = 4:11,
entropy = TRUE,
repetitions = 10,
project = "new")
```

Loading the product of this function and plotting cross entropy for each value of K (number of ancestry components). People often pick the K with the lowest cross-entropy, but we see a kind of asymptote.

```{r}

project = readRDS("jap_structure.rds")

plot(project, col = "blue", pch = 19, cex = 1.2)

```
K = 8 recaptures the population structure inferred in Gutaker et al. 2020 and Alam et al. 2021. Don't run this.

```{r}

K = 8

best = which.min(cross.entropy(project, K = K))

admix = as.data.frame(Q(project , K = K, run = best))
admix$ID = kept$ID

tai_meta = read.csv("taiwan_metadata.csv")
tai_meta = tai_meta %>% dplyr::select(ID,K11)
admix = left_join(admix,tai_meta) 

admix = admix %>% dplyr::arrange(K11)

admix_long = pivot_longer(admix,cols=1:K)
admix_long = admix_long %>% rowwise() %>%
  dplyr::mutate(ord_ID = paste0(ID,"_",name))


ggplot(admix_long,aes(x=factor(ID),y = value,color = factor(name))) +
  geom_col() +
  facet_grid(~K11,space = "free",scales = "free") +
  theme(axis.text.x = element_blank())

```

Before we run the actual GEA, looking at the correlatedness between the three environmental variables:


```{r}

# env_list is the same as env_variables, but is a list of separate dataframes for each condition
env_list = readRDS("env_list.rds")

cor(env_list[["present"]][,4:6])

```

Performing PCA

```{r}
pred.pca <- vegan::rda(env_list[["present"]][,4:6], scale=T)

summary(pred.pca)$cont
```
```{r}
screeplot(pred.pca, main = "Screeplot: Eigenvalues of Predictor Variables")

```

PC1 seems largely explanatory, using that as the environmental variable for GEA

```{r}
env.pc1 = vegan::scores(pred.pca, choices=1, display="sites", scaling=0)
#saveRDS(env.pc1,"env.PC1.rds")
```

Looking at correlation between PC1 and the environmental variables

```{r}
cor(data.frame(env_list[["present"]][,4:6],PC1 = env.pc1))
```

Ran the following (across different values of K on the HPC)

```{r}
#lf_matrix = data.table::fread("japonica_filtered_maf0.05.lfmm")
#lf_PC_K8 = lfmm2(input = lf_matrix, env = env.pc1,
#      K = 8)

#p_PC_K8 = lfmm2.test(lf_PC_K8,lf_matrix,env.pc1)
```

Looking at distribution of p-values

```{r}

p_PC_K8 = readRDS("p_PC1_K8.rds")

hist(p_PC_K8$pvalues)

```
qqplot is a bit inflated, but does imply presence of outliers:

```{r}

qqplot(rexp(length(p_PC_K8$pvalues), rate = log(10)),
       -log10(p_PC_K8$pvalues), xlab = "Expected quantile",
       pch = 19, cex = .4)
abline(0,1)
```
(This kind of qqplot is not completely unheard of:
https://bcm-uga.github.io/pcadapt/articles/pcadapt.html)


Importing SNP positions:

```{r}
snp_table = read.table("japonica_filtered_maf0.05.map")

snp_table$p = p_PC_K8$pvalues

snp_table$index = 1:length(snp_table$p)
```


Correcting for false-positives and making a Manhattan plot

```{r}

fdr_K8 = p.adjust(p_PC_K8$pvalues,"BY",length(p_PC_K8$pvalues))

sum(fdr_K8 < 0.1)

snp_table$fp = fdr_K8 

snp_table = snp_table %>% dplyr::mutate(sig = case_when(fp < 0.1 ~ "sig.",
                                            fp >= 0.1 ~ "non-sig."
))


ggplot(snp_table,aes(x = index, y = -log10(p),color=sig)) +
  geom_point()
```



This is arguably a generous cutoff, but I think having more candidate loci makes the genetic offset calculation more robust, or at least not driven by one or two loci.

A lot of the SNPs are clearly linked. Picking the most significant SNPs from every contiguous series of SNPs (within 100kb) and making another Manhattan plot:


```{r}
snps_K8 = snp_table %>% dplyr::filter(fp < 0.1)

chain = rep(1,length(snps_K8$V4))

for (i in 2:length(snps_K8$V4)){
  if (snps_K8$V1[i] != snps_K8$V1[i-1] | 
      snps_K8$V4[i] - snps_K8$V4[i-1] > 100000){
    chain[i:length(chain)] = chain[i] + 1
  }
}

snps_K8$chain = chain

snps_K8_loci = snps_K8 %>% dplyr::group_by(chain) %>%
  dplyr::filter(p == min(p)) %>%
  dplyr::filter(V4 == min(V4))

#Show these loci on Manhattan plot:

ggplot(snp_table,aes(x = index, y = -log10(p))) +
geom_point(color = "grey") +
geom_point(data = snps_K8_loci, mapping = 
             aes(x = index, y = -log10(p)),color = "blue")


```
Calculated genetic offset using the following code on the hpc (don't run):


```{r}
offset_2.6 = genetic.gap(input = lf_matrix,
            env = env_list[["present"]][,4:6],
            pred.env = env_list[["SSP1-2.6"]][,4:6],
            candidate.loci = snps_K8_loci$index,
            K = 8
            )

offset_7.0 = genetic.gap(input = lf_matrix,
            env = env_list[["present"]][,4:6],
            pred.env = env_list[["SSP3-7.0"]][,4:6],
            candidate.loci = snps_K8_loci$index,
            K = 8
            )

offset_8.5 = genetic.gap(input = lf_matrix,
            env = env_list[["present"]][,4:6],
            pred.env = env_list[["SSP5-8.5"]][,4:6],
            candidate.loci = snps_K8_loci$index,
            K = 8
            )


offset_past = genetic.gap(input = lf_matrix,
            env = env_list[["present"]][,4:6],
            pred.env = env_list[["past"]][,4:6],
            candidate.loci = snps_K8_loci$index,
            K = 8
            )

```

Importing these objects:

```{r}
offset_2.6 = readRDS("offset_2.6_PC1_K8.rds")
offset_7.0 = readRDS("offset_7.0_PC1_K8.rds")
offset_8.5 = readRDS("offset_8.5_PC1_K8.rds")
offset_past = readRDS("offset_past_PC1_K8.rds")
```

Pulling the offset statistics into a dataframe with locations of the accessions


```{r}

# this is the order of the japonica accessions in my genotype matrix
japs = read.table("japs_order.txt")

japs_offset = data.frame(japs$V1,
                  offset_2.6$offset,offset_2.6$distance,
                  offset_7.0$offset,offset_7.0$distance,
                  offset_8.5$offset,offset_8.5$distance,
                  offset_past$offset,offset_past$distance)


colnames(japs_offset) = c("ID",
                   "offset_2.6","rona_2.6",
                   "offset_7.0","rona_7.0",
                   "offset_8.5","rona_8.5",
                   "offset_past","rona_past"
                   )


japs_offset_meta = dplyr::left_join(japs_offset,japs_geotagged)
```

Looking at the maximum offset value for plotting range

```{r}
max(japs_offset_meta$offset_2.6,
    japs_offset_meta$offset_7.0,
    japs_offset_meta$offset_8.5,
    japs_offset_meta$offset_past)
```


Making the plots


```{r}
plot_off_2.6 = ggplot(asia) +
  geom_sf() +
  geom_point(data = japs_offset_meta, 
             mapping = aes(x = LON,y = LAT, color = offset_2.6)) +
            scale_colour_gradientn( colors = c("lightblue1","darkblue"),limits=c(0,0.1)) +
   theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_blank(),
        legend.position = "none",
        axis.ticks = element_blank()) +
  ggtitle("Model 2.6")
  
```

```{r}
plot_off_7.0 = ggplot(asia) +
  geom_sf() +
  geom_point(data = japs_offset_meta, 
             mapping = aes(x = LON,y = LAT, color = offset_7.0)) +
            scale_colour_gradientn( colors = c("lightblue1","darkblue"),limits=c(0,0.1)) +
   theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_blank(),
        legend.position = "none",
        axis.ticks = element_blank()) +
  ggtitle("Model 7.0")
  
```


```{r}

plot_off_8.5 = ggplot(asia) +
  geom_sf() +
  geom_point(data = japs_offset_meta, 
             mapping = aes(x = LON,y = LAT, color = offset_8.5)) +
   scale_colour_gradientn( colors = c("lightblue1","darkblue"),limits=c(0,0.1)) +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_blank(),
        legend.position = "none",
        axis.ticks = element_blank()) +
  ggtitle("Model 8.5")
```


```{r}
plot_off_past = ggplot(asia) +
  geom_sf() +
  geom_point(data = japs_offset_meta, 
             mapping = aes(x = LON,y = LAT, color = offset_past)) +
            scale_colour_gradientn( colors = c("lightblue1","darkblue"),limits=c(0,0.1)) +
   theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_blank(),
        legend.position = "none",
        axis.ticks = element_blank()) +
  ggtitle("Past")
  
```

This is just extracting the legend from one of the plots

```{r}

plot_with_legend = ggplot(asia) +
  geom_sf() +
  geom_point(data = japs_offset_meta, 
             mapping = aes(x = LON,y = LAT, color = offset_8.5)) +
   scale_colour_gradientn( "Offset",colors = c("lightblue1","darkblue"),limits=c(0,0.1)) +
  theme(axis.text = element_blank(),
        axis.title = element_blank(),
        panel.background = element_blank(),
        axis.ticks = element_blank()) +
  ggtitle("Model 8.5")

legend <- cowplot::get_legend(
  plot_with_legend +  
      theme(legend.position = "bottom") 
)
```

Plotting them all together:

```{r}
prow = cowplot::plot_grid(plot_off_2.6,plot_off_7.0,
                          plot_off_8.5,plot_off_past)

cowplot::plot_grid(prow,legend,ncol = 1,rel_heights = c(7,1))
```

